The Gaia-ESO Survey: Combined estimates from multiple stellar analysis pipelines

### 2016-10-14 article/ms.tex
Add some paragraphs about technical flags

@@

Flag propagation.

@@

We list the nomenclature and descriptions for all model parameters in Table 
\ref{tab:model-parameters}. The number of total model parameters is given by
\begin{eqnarray}
    N_{parameters} = N_{calibrators} + N_{missing} + 6N_{nodes} + \frac{N_{nodes}!}{2(N_{nodes} - 2)!}
\end{eqnarray}
where $N_{calibrators}$ accounts for the true values of the calibrators, the
$N_{missing}$ models the missing data points for the calibrator spectra, the
$6N_{nodes}$ summarises parameters that model biases, as well as systematic
and random uncertainties. The final term $\frac{N_{nodes}!}{2(N_{nodes} - 2)!}$
is the number of correlation coefficients between all nodes.\footnote{
This inference problem is numerically unstable, as there are many model parameters
which are used toi

 construct a covariance matrix that is sampled

In practice
we include the Cholesky decomposition $L$ of the $N \times N$ correlation matrix $\rho$
as model parameters instead of directly modeling the $\rho$ terms.  The correlation
matrix can be reconstructed by $\rho = (IL)\cdot(IL)^T$ where $I$ is the $N \times N$
identity matrix. This parameterisation helps ensure that the resulting covariance
matrix $\mathbf{\Sigma}$ is positive semi-definite, and can therefore be used for
sampling.}

>>>>

\subsection{Technical flags}
\label{sec:technical-flags}

In the fifth \gaiaeso\ data release we required all analysis nodes to either
supply results for a spectrum, or to provide a technical flag (column \texttt{TECH})
explaining no results were reported. We list all relevant technical flags and their
interpretation in Appendix \ref{app:flags}. There is a large variation in specificity 
between flags. For example, some flags indicate that an analysis pipeline did not 
converge (for whatever reason), whereas some flags indicate a result is uncertain 
due to a data reduction issue producing a `picket-fence'-like pattern in the 
spectrum. The former is provided automatically and may reflect uncertainty 
\emph{only} in the results from the node who reported the flag. However, the latter 
example was presumably flagged because a human looked at the data, and this flag
should raise uncertainty in \emph{all} results using the same spectrum.  Similar 
examples can be drawn for flags describing issues intrinsic to the star (e.g., high
rotational velocity), and not just a single spectrum.  For these reasons, there are 
situations where technical flags ought to be propagated to other results derived from 
the same spectrum, or other results reported for the same star.

We investigated usage patterns between nodes to evaluate whether any flag propagation
was warranted. A heatmap of common flag usage for WG11 nodes is shown in Figure
\ref{fig:wg11-flag-heatmap}.
% TODO: Include heat maps of other WGs?
If all nodes reported the same flags under common conditions, then the node-to-node 
segments in Figure \ref{fig:wg11-flag-heatmap} would show 1-to-1 correlations. This is 
not the case; we find the flag usage to be mostly uncorrelated between different nodes. 
When one node reports a \emph{specific} flag indicating an issue that is intrinsic to 
the spectrum or star, it is highly unlikely that \emph{any} other node has reported the 
same problem.\footnote{In some sense, this is a consequence of having `many eyes on the 
data'.} The most significant correlations are seen in the usage of non-descript flags,
where nodes will report that the derived stellar parameters are outside of the range 
where they consider their results to be `physically realistic' (e.g., outside of a grid
or the effective temperature $\teff$ is too high or low to be reliable).

Based on our examination of technical flag usage between nodes, we assigned every node 
result with a boolean flag indicating whether the supplied node results ($\teff$, $\logg$,
$\feh$) should be used in the homogenisation process.\footnote{This flag is only present
in an internal PostgreSQL database, which also includes the propagated flag explaining
why a record should not be used in the homogenisation process, and which record the flag
originated from.}  We did not use any results that were marked with any of the technical
flags listed in Table \ref{tab:wg11-node-specific-flags-failing-qc}.  We further discarded 
results flagged as having $v\sin{i} > 20$~km~s$^{-1}$ (flags 11020 and 11050) unless the 
reporting node was OACT, as our examination (over multiple data releases) revealed that 
the OACT node can properly account for fast rotating stars, whereas some nodes may not 
even detect the presence of significant rotation.

We discarded results from IACAIP that were flagged to be an incomplete spectrum or 
missing wavelengths (flag 10105), as IACAIP seemingly used this flag extensively to 
indicate \emph{any} problem with their own analysis.  For all other nodes that report
flag 10105, we propagated this flag to all results derived from the same spectrum,
and discarded those node estimates.  However, if \emph{any} node reported one of the
data reduction issues in Table \ref{tab:wg11-spectrum-specific-flags-failing-qc}
(e.g., saturated or broken spectrum, poor sky subtraction) then we propagated this 
flag to other results derived from the same spectrum, and discarded those results.
For stars with multiple observations, we were also able to identify situations where 
all nodes reported significantly different stellar parameters, all in the same 
systematic direction (e.g., $\teff$ values reported by all nodes were $\sim$300~K for 
one spectrum). In all cases, these systematic differences arose from archival spectra. 
A visual inspection of these spectra revealed substantial data reduction issues, and 
therefore we discarded all results derived from these problematic spectra.  Finally, 
we discarded \emph{all} node results for a star (matched by \texttt{CNAME}) if any of 
the spectra were flagged -- by any node -- as being a suspected multiple stellar system 
(10320, 13027, or 20020). The total fraction of WG11 records that were discarded due to
flags, including propagated flags, is 23.7\% (11,553 of 48,720), however only 580 unique 
stars analysed by WG11 had all results discarded due to flag usage (of 4,941; 11.7\%).

%FIGURE: wg11-flag-heatmap: Show 2D matrix heatmap of flag usage by WG11 nodes

%TABLE IN APPENDIX \label{app:flags}: <no-name>: List of all relevant flags and their meaning.

%TABLE IN APPENDIX \label{app:flags}: wg11-node-specific-flags-failing-qc: List of node-specific flags that caused us to discard results. Note in the caption the conditions for OACT & vsini, as well as IACAIP with missing wavelengths.

%TABLE IN APPEDIX \label{app:flags}: wg11-spectrum-specific-flags-failing-qc: List of spectrum-specific flags that caused us to discard results. Note in the caption that we did something unusual for IACAIP.

@@

@@

We list the nomenclature and descriptions for all model parameters in Table 
\ref{tab:model-parameters}. The number of total model parameters is given by
\begin{eqnarray}
    N_{parameters} = N_{calibrators} + N_{missing} + 6N_{nodes} + \frac{N_{nodes}!}{2(N_{nodes} - 2)!}
\end{eqnarray}
where $N_{calibrators}$ accounts for the true values of the calibrators, the
$N_{missing}$ models the missing data points for the calibrator spectra, the
$6N_{nodes}$ summarises parameters that model biases, as well as systematic
and random uncertainties. The final term $\frac{N_{nodes}!}{2(N_{nodes} - 2)!}$
is the number of correlation coefficients between all nodes.\footnote{
This inference problem is numerically unstable, as there are many model parameters
AND TODO
we include the Cholesky decomposition $L$ of the $N \times N$ correlation matrix $\rho$
as model parameters instead of directly modeling the $\rho$ terms.  The correlation
matrix can be reconstructed by $\rho = (IL)\cdot(IL)^T$ where $I$ is the $N \times N$
identity matrix. This parameterisation helps ensure that the resulting covariance
matrix $\mathbf{\Sigma}$ is positive semi-definite, and can therefore be used for
sampling.}
